---
title: "CRITICAL COASTAL HABITAT ASSESSMENT"
output: 
  flexdashboard::flex_dashboard:
    logo: www/tarponlogo.png
    social: menu
    source_code: "https://github.com/tbep-tech/ccha-workflow"
runtime: shiny
css: styles.css
---
  
```{r setup, include=F}
knitr::opts_chunk$set(echo = F, warning = F, message = F)

library(here)
library(ggplot2)
library(tidyr)
library(dplyr)
library(ggridges)
library(reactable)
library(shiny)
library(lubridate)
library(stringr)
library(forcats)
library(htmltools)
library(htmlwidgets)
library(shinyWidgets)

source(here('R/funcs.R'))

load(file = here('data/vegdat.RData'))
load(file = here('data/treedat.RData'))

thm <- theme_minimal(base_size = 20) + 
  theme(
    panel.grid.minor = element_blank() 
  )

# initial options
sitopt <- unique(vegdat$site)
sppopt <- unique(vegdat$species) %>% 
  sort %>% 
  .[!. %in% c('Boardwalk', 'none/detritus', 'Open Water', 'Unknown')]

# # style file
# styles <- readLines('https://raw.githubusercontent.com/tbep-tech/css-styling/master/styles.css')
# writeLines(styles, 'styles.css')
```

```{r reactives}
# phase by site/zone
sitplo1 <- reactive({
  
  # input
  sitsel1 <- input$sitsel1
  zonsel1 <- input$zonsel1
  
  req(zonsel1)
  
  out <- sitezonedst_plo1(vegdat, site = sitsel1, zonefct = zonsel1, thm)
  
  return(out)
  
})

# distance by phase/site
sitplo2 <- reactive({
  
  # input
  sitsel1 <- input$sitsel1
  zonsel1 <- input$zonsel1
  
  req(zonsel1)
  
  out <- sitezonedst_plo2(vegdat, site = sitsel1, zonefct = zonsel1, thm)
  
  return(out)
  
})

# tabular vegetation summary by site
vegtab1 <- reactive({
  
  # input
  sitsel2 <- input$sitsel2
  zonsel2 <- input$zonsel2
  varsel1 <- input$varsel1

  req(zonsel2)
  
  out <- sitezonesum_tab(vegdat, sitsel2, zonsel2, varsel1)

  return(out)
  
})

# plot vegetation summary by site
vegplo1 <- reactive({
  
  # input
  sitsel2 <- input$sitsel2
  zonsel2 <- input$zonsel2
  varsel1 <- input$varsel1
  topsel <- input$topsel
  metsel <- input$metsel

  req(zonsel2)
  
  out <- sitesum_plo(vegdat, sitsel2, delim = metsel, top = topsel, var = varsel1, zonefct = zonsel2, thm = thm)
  
  return(out)
  
})

# summary plot by species, all sites
vegplo2 <- reactive({
  
  # input
  sitsel2 <- input$sitsel2
  zonsel2 <- input$zonsel2
  
  req(zonsel2)
  
  out <- sitezonedst_plo3(vegdat, sitsel2, zonefct = zonsel2, thm = thm)
  
  return(out)
  
})

# tabular tree summary by site
treetab1 <- reactive({
  
  # input
  sitsel3 <- input$sitsel3
  zonsel3 <- input$zonsel3
  varsel2 <- input$varsel2
  aggsel <- input$aggsel

  req(zonsel3)
  
  out <- treesum_tab(treedat, sitsel3, byspecies = aggsel, zonefct = zonsel3, var = varsel2)
  
  return(out)
  
})

# plot tree summary by site
treeplo1 <- reactive({
  
  # input
  sitsel3 <- input$sitsel3
  zonsel3 <- input$zonsel3
  varsel2 <- input$varsel2
  aggsel <- input$aggsel

  req(zonsel3)
  
  out <- treesum_plo(treedat, sitsel3, byspecies = aggsel, zonefct = zonsel3, var = varsel2, thm = thm)
  
  return(out)
  
})

# summary plot by species, all sites
allplo1 <- reactive({
  
  # input
  sppsel <- input$sppsel
  stesel <- input$stesel
  varsel3 <- input$varsel3
  
  req(stesel)
  
  out <- sppsum_plo(vegdat, sppsel, var = varsel3, sitefct = stesel, thm)
  
  return(out)
  
})
```

1 ZONES BY SITE {data-navmenu="PAGE SELECTION"}
=======================================================================
  
Column {data-width=200 .tabset .tabset-fade}
-----------------------------------------------------------------------
  
### Site selection

```{r}
column(12,  
  selectInput('sitsel1', 'Select site:', choices = sitopt, width = '100%')
)
column(12, 
  renderUI({
    
    # input
    sitsel1 <- input$sitsel1
    
    tosel <- vegdat %>% 
      filter(site %in% !!sitsel1) %>% 
      unite('zonefct', zone, zone_name, sep = ': ') %>% 
      mutate(zonefct = factor(zonefct, levels = sort(unique(zonefct)))) %>% 
      pull(zonefct) %>% 
      levels
    
    selectInput('zonsel1', 'Select zone:', choices = tosel, selectize = T, multiple = T, selected = tosel, width = '100%')
    
  })
)
```

### About

Column {.tabset .tabset-fade data-width=700}
-----------------------------------------------------------------------

### Zones by phase

```{r}
output$sitplo1 <- renderPlot(sitplo1())
plotOutput('sitplo1')
```

### Zone length by phase

```{r}
output$sitplo2 <- renderPlot(sitplo2())
plotOutput('sitplo2')
```

2 VEGETATION SURVEYS {data-navmenu="PAGE SELECTION"}
=======================================================================
  
Column {data-width=200 .tabset .tabset-fade}
-----------------------------------------------------------------------
  
### Summary options

```{r}
column(12,  
  selectInput('sitsel2', 'Select site:', choices = sitopt, width = '100%')
)
column(12, 
  renderUI({
    
    # input
    sitsel2 <- input$sitsel2
    
    tosel <- vegdat %>% 
      filter(site %in% !!sitsel2) %>% 
      unite('zonefct', zone, zone_name, sep = ': ') %>% 
      mutate(zonefct = factor(zonefct, levels = sort(unique(zonefct)))) %>% 
      pull(zonefct) %>% 
      levels
    
    selectInput('zonsel2', 'Select zone:', choices = tosel, selectize = T, multiple = T, selected = tosel, width = '100%')
    
  })
)
column(12, 
  selectInput('varsel1', 'Select summary variable:', choices = list('Freq. Occ (%)' = 'fo', 'Basal cover (%)' = 'cover'), width = '100%')
)
```

### About

Column {.tabset .tabset-fade data-width=700}
-----------------------------------------------------------------------

### Tabular summary

```{r}
output$vegtab1 <- renderReactable(vegtab1())
reactableOutput('vegtab1')
```

### Species summary

```{r}
output$vegplo1 <- renderPlot(vegplo1())
fillCol(flex = c(NA, 1),
  column(12, 
    column(6, sliderInput('topsel', 'Select number of species:', min = 1, max = 30, step = 1, value = 10, width = '100%')),
    column(6, sliderInput('metsel', 'Select number of meter bins:', min  = 1, max = 30, step = 1, value = 10, width = '100%'))
  ),
  plotOutput('vegplo1')
)
```

### All species

```{r}
output$vegplo2 <- renderPlot(vegplo2())
plotOutput('vegplo2')
```

3 TREE SURVEYS {data-navmenu="PAGE SELECTION"}
=======================================================================
  
Column {data-width=200 .tabset .tabset-fade}
-----------------------------------------------------------------------
  
### Summary options

```{r}
column(12,  
  selectInput('sitsel3', 'Select site:', choices = sitopt, width = '100%')
)
column(12, 
  renderUI({
    
    # input
    sitsel3 <- input$sitsel3
    
    tosel <- treedat %>% 
      filter(site %in% !!sitsel3) %>% 
      unite('zonefct', zone, zone_name, sep = ': ') %>% 
      mutate(zonefct = factor(zonefct, levels = sort(unique(zonefct)))) %>% 
      pull(zonefct)
    
    selectInput('zonsel3', 'Select zone:', choices = tosel, selectize = T, multiple = T, selected = tosel, width = '100%')
    
  })
)
column(12, 
  selectInput('varsel2', 'Select summary variable:', 
    choices = list(
      'Absolute species density (trees/m2)' = 'trees_m2',
      'Absolute species density (trees/ha)'= 'trees_ha', 
      'Species average basal area (cm2/m2)' =  'cm2_m2',
      'Species absolute cover (m2/ha)' = 'm2_ha', 
      'Relative cover (%)' = 'relcov_per'
    ),
    width = '100%'
  )
)
column(12, 
  materialSwitch(inputId = 'aggsel', label = 'Show results by species?', value = T, width = '100%')
)
```

### About

Column {.tabset .tabset-fade data-width=700}
-----------------------------------------------------------------------

### Tabular summary

```{r}
output$treetab1 <- renderReactable(treetab1())
reactableOutput('treetab1')
```

### Species by distance

```{r}
output$treeplo1 <- renderPlot(treeplo1())
plotOutput('treeplo1')
```

4 BETWEEN SITE COMPARISONS {data-navmenu="PAGE SELECTION"}
=======================================================================
  
Column {data-width=200 .tabset .tabset-fade}
-----------------------------------------------------------------------
  
### Summary options

Summaries are for vegetation data only. 

```{r}
column(12, 
  selectInput('sppsel', 'Select species:', choices = sppopt, selected = 'Avicennia germinans', width = '100%')
)
column(12, 
  renderUI({
    
    # input
    sppsel <- input$sppsel
    
    tosel <- vegdat %>% 
      filter(species %in% !!sppsel) %>% 
      pull(site) %>% 
      unique %>% 
      sort
    
    selectInput('stesel', 'Select site:', choices = tosel, selectize = T, multiple = T, selected = tosel, width = '100%')
    
  })
)
column(12, 
  selectInput('varsel3', 'Select summary variable:', choices = list('Freq. Occ (%)' = 'fo', 'Basal cover (%)' = 'cover'), width = '100%')
)
```

### About

Column {.tabset .tabset-fade data-width=700}
-----------------------------------------------------------------------

### Species by zone

```{r}
output$allplo1 <- renderPlot(allplo1())
plotOutput('allplo1')
```
