---
title: CCHA site assessments
author: "MW Beck"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    code_folding: hide
---

```{r setup, echo = T, message = F, warning = F}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)

box::use(
  here[...], 
  ggplot2[...], 
  tidyr[...],
  dplyr[...], 
  ggridges[...]
)

load(file = here('data/cchadat.RData'))

thm <- theme_minimal(base_size = 14) + 
  theme(
    panel.grid.minor = element_blank() 
  )
```

Plot showing number of times a zone was identified across transects, dates.

```{r, fig.height = 5, fig.width = 7}
toplo <- cchadat %>%
  select(site, sample, zone_name) %>% 
  unique %>% 
  filter(!is.na(zone_name)) %>% 
  group_by(zone_name) %>% 
  summarize(
    cnt = n(), 
    .groups = 'drop'
  ) %>% 
  arrange(cnt) %>% 
  mutate(
    zone_name = factor(zone_name, levels = zone_name)
  )

ggplot(toplo, aes(y = zone_name, x = cnt)) + 
  geom_bar(stat = 'identity', alpha = 0.7) + 
  thm +
  theme(
    axis.text.y = element_text(size = 8)
  ) +
  scale_x_continuous(expand = c(0, 0), breaks = seq(0, max(toplo$cnt))) +
  labs(
    y = NULL, 
    x = 'Unique counts across sites, dates'
  )
```

Plot showing total distance zone was recorded aross transect, dates.

```{r, fig.height = 5, fig.width = 7}
toplo <- cchadat %>%
  select(site, sample, zone_name, meter) %>% 
  unique %>% 
  filter(!is.na(zone_name)) %>% 
  group_by(site, sample, zone_name) %>% 
  summarize(
    dist = max(meter) - min(meter),
    .groups = 'drop'
  ) %>% 
  group_by(zone_name) %>% 
  summarise(
    dist = sum(dist, na.rm = T)
  ) %>% 
  arrange(dist) %>% 
  mutate(
    zone_name = factor(zone_name, levels = zone_name)
  )
  
ggplot(toplo, aes(y = zone_name, x = dist)) + 
  geom_bar(stat = 'identity', alpha = 0.7) + 
  thm +
  theme(
    axis.text.y = element_text(size = 8)
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  labs(
    y = NULL, 
    x = 'Total distance recorded, all dates'
  )
```

An example showing zones for two sample dates at one site. 

```{r, fig.height = 3, fig.width = 12}
site <- 'Big Bend - TECO'
toplo <- cchadat %>% 
  filter(site == !!site) %>% 
  select(zone_name, zone, sample, meter, date) %>% 
  unique %>% 
  unite('sample', sample, date, sep = ': ') %>% 
  mutate(
    sample = factor(sample, levels = rev(unique(sample)))
  ) %>% 
  group_by(sample, zone_name)

levs <- toplo %>% 
  filter(grepl('^one', sample)) %>%
  select(zone_name, zone) %>% 
  unique %>% 
  filter(!duplicated(zone)) %>% 
  pull(zone_name)

toplo <- toplo %>% 
  mutate(
    zone_name = factor(zone_name, levels = levs)
  )

ggplot(toplo, aes(x = meter, y = sample)) + 
  geom_line(aes(color = zone_name), stat = 'identity', lwd = 20, alpha = 0.7) + 
  scale_x_continuous(breaks = seq(0, max(toplo$meter), by = 10)) +
  guides(color = guide_legend(override.aes = list(lwd = 7))) +
  thm +
  theme(
    panel.grid.major.y = element_blank()
  ) +
  labs(
    x = 'Meters', 
    y = NULL, 
    subtitle = paste0('Site: ', site),
    color = 'Zone name'
  )
```

The same site, but showing species data along the transect.

```{r fig.height = 8, fig.width = 12}
site <- 'Big Bend - TECO'

tofilt <- c('Unknown', 'Open Water', 'none/detritus', 'Woody Debris')

toplo <- cchadat %>% 
  filter(site == !!site) %>% 
  filter(!species %in% tofilt) %>% 
  mutate(
    zone_name = factor(zone_name, levels = levs)
  ) %>% 
  unite('sample', sample, date, sep = ': ')

ggplot(toplo, aes(x = meter, y = species, height = pcent_basal_cover / 100, fill = zone_name, color = zone_name)) + 
  geom_ridgeline(stat = 'identity') + 
  scale_x_continuous(expand = c(0, 0), breaks = seq(0, max(toplo$meter), by = 20)) + 
  thm + 
  # theme(panel.grid.major.y = element_blank()) + 
  labs(
    x = 'Meters', 
    color = 'Zone name', 
    fill = 'Zone name', 
    y = 'Species relative basal cover', 
    subtitle = paste0('Site: ', site)
  ) + 
  facet_wrap(~sample, ncol = 2)
```

To evaluate simple changes in zones over time, an estimate of total distance per zone between time points could be evaluated.

```{r}
site <- 'Big Bend - TECO'

toplo <- cchadat %>% 
  filter(site == !!site) %>% 
  select(zone_name, zone, sample, meter, date) %>% 
  unique %>% 
  unite('sample', sample, date, sep = ': ') %>% 
  group_by(zone_name, zone, sample) %>% 
  summarize(
    dist = max(meter) - min(meter), 
    .groups = 'drop'
  )
```

